Association between Hormone and metabolites
========================================================

First, type in the console: opts_knit$set(root.dir = "...") where the data file is located.

1. Import hormone data
-------------------------------
```{r}
load("data.RData")
hormone = read.csv("data/Hormone data_sul.csv",stringsAsFactors = F)
colnames(hormone)[10:12] = c("Melatonin", "Cortisol", "Estradiol")
```

```{r}
hormone = hormone[order(hormone$SW_Nr, hormone$Probennahme_Dat, hormone$Proben_Nr),]
hormone$Probennahme_Dat=as.character(hormone$Probennahme_Dat)
```


2. Categorize the sampling time, calculate the aggregated hormone levels and metabolite concentrations at different time intervals 
-------------------------------
### 2.1 set the samplying time intervals (0:00~3:00, 3:00-6:00, 6:00~9:00, 9:00:00~ 12:00:00, 12:00~15:00, 15:00~18:00, 18:00~21:00, 21:00~23:59)
```{r}
hormone$hour = strptime(hormone$Probennahme_Uhr, "%H:%M:%S")
hormone$Probennahme_Dat = strptime(hormone$Probennahme_Dat, "%Y.%m.%d")
hormone$Probennahme_Uhr = as.character(hormone$Probennahme_Uhr)
hormone$Probennahme_Uhr = strptime(hormone$Probennahme_Uhr, "%H:%M")
hormone$hour_cut = cut(hormone$hour, breaks = strptime((c("0:00:00", "3:00:00", "6:00:00", "9:00:00", "12:00:00","15:00:00","18:00:00","21:00:00","23:59:59"), "%H:%M:%S"))

data.merged$hour = strptime(data.merged$Probennahme_Uhr, "%H:%M")
data.merged$hour_cut = cut(data.merged$hour, 
                           breaks = strptime(c("0:00:00", "3:00:00", "6:00:00", "9:00:00", "12:00:00","15:00:00","18:00:00","21:00:00","23:59:59"), "%H:%M:%S"))
metabo = data.merged

```

### 2.2 Aggregate the metabolite concentration according to the time intervales
```{r}
metabo[, valid_measures] = scale(log(metabo[, valid_measures]))
metabo_aggre = aggregate(metabo[,valid_measures], by=list(metabo$SW_Nr, metabo$hour_cut,metabo$Probennahme_Dat, metabo$Schichtdienst), mean, na.rm=T)
colnames(metabo_aggre)[1:4]=c("SW_Nr","hour", "date", "shift")

hormone[, c("Melatonin", "Cortisol", "Estradiol")] = scale(log(hormone[, c("Melatonin", "Cortisol", "Estradiol")]))
hormone_aggre = aggregate(hormone[, c("Melatonin","Cortisol","Estradiol")], by=list(hormone$SW_Nr, hormone$hour_cut, hormone$Probennahme_Dat, hormone$Schichtdienst), mean, na.rm=T)
colnames(hormone_aggre)[1:4]=c("SW_Nr","hour", "date", "shift")
```

### 2.3 Match the metabolite samples with the hormone samples
```{r}
match_metabo_hormone = merge(metabo_aggre, hormone_aggre)
match_metabo_hormone = match_metabo_hormone[order(match_metabo_hormone$SW_Nr, match_metabo_hormone$date, match_metabo_hormone$hour),]
```

3. Correlation analysis between metabolite concentrations and hormone levels.
------------------------------
```{r comment='heatmap of correlation between metabolites and hormorns'}
require(corrplot)
getwd()
pdf("TEMP/correlation between metabolite with hormones.pdf", width = 20, height=20)
#col1 = colorRampPalette(c("#053061","#2166AC","#4393C3","#92C5DE", "#D1E5F0", "#FFFFFF", "#FDDBC7","#F4A582","#D6604D", "#B2182B",  "#67001F"))  
corrplot(cor(match_metabo_hormone[,-c(1:4)], use="pair"), tl.col = "black")
dev.off()
```

4. Analyze the metabolite secretion rate in categorized time intervals.
------------------------------
### 4.1 Estimate the association between hormone and metabolites in linear mixed effect model
```{r}
require(nlme)
rst = NULL
for(i in valid_measures){
  match_metabo_hormone$m = match_metabo_hormone[,i]
  model = lme(m~ Melatonin + Cortisol + Estradiol, data = match_metabo_hormone, random = ~1|SW_Nr, na.action = na.omit)
  tmp = summary(model)$tTable
  rst = rbind(rst, c(tmp[1,], tmp[2, ], tmp[3,]))
}
rownames(rst) = valid_measures
write.csv(rst, "TEMP/association between metabolites and hormones_lme.csv")
```

### 4.2 Analyze the metabolite secretion rate in categorized time intervals.
#### 4.2.1 Adding personal characteristics information to the metabolite dataset
```{r}
metabo_aggre = merge(metabo_aggre, samples.addition, by.x = "SW_Nr", by.y = "P_ID")
metabo_aggre$date = strptime(metabo_aggre$date, "%Y.%m.%d")
metabo_aggre$hour = strptime(metabo_aggre$hour, "%Y-%m-%d %H:%M:%S")
metabo_aggre = metabo_aggre[order(metabo_aggre$SW_Nr, metabo_aggre$date, metabo_aggre$hour),]

metabo_aggre$hour2 = metabo_aggre$hour
for(p in levels(metabo_aggre$SW_Nr)){
  for(sh in levels(metabo_aggre$shift)){
    subset = which(metabo_aggre$SW_Nr==p& metabo_aggre$shift==sh)
    a = metabo_aggre[subset, 3]
    metabo_aggre$hour2[subset] = metabo_aggre$hour2[subset]+as.numeric(a-a[1])
  }
}
metabo_aggre$hour2 = as.character(metabo_aggre$hour2)

metabo_aggre$date2 = sapply(metabo_aggre$hour2, function(x) strsplit(x, split = " ")[[1]][1])
metabo_aggre$date2 = as.factor(metabo_aggre$date2)
levels(metabo_aggre$date2) = c("day1","day2", "day3", "day4", "day5")

levels(metabo_aggre$Kontrolle)=c("Night", "Day")
```

#### 4.2.2 Investigate the differences of metabolite profile between case and control group in the day shift work
```{r warning=FALSE, message=FALSE, results='hide'}
require(gee)
rst = NULL

for(i in valid_measures){
  metabo_aggre$m = metabo_aggre[,i]
   model = suppressMessages(
      gee(m ~ as.factor(Kontrolle)+ Alter + BMI + as.factor(AR_Rauch_zurzt)+as.factor(SD),
                id = SW_Nr, 
                data = metabo_aggre, 
                subset = shift=="Tagschicht"& SW_Nr!="SW1041"& Alter>45,
                na.action=na.omit, 
                corstr = "exchangeable"
                  )
      )
    rst = rbind(rst, summary(model)$coef[2,])
}
rownames(rst) = valid_measures
rst = data.frame(rst, p.value = 2*pnorm(-abs(rst[,5])))
rst = data.frame(rst, fdr = p.adjust(rst$p.value, method = "BH"), bonf = p.adjust(rst$p.value, method = "bonf"))
rst$Estimate = -rst$Estimate
write.csv(rst, "TEMP/categorized time_3h_Chronic effect of night shift work_GEE_daywork_age_BMI_smoking_thyroid_exclude diab_over 45.csv")

```

#### 4.2.3 Investigate the differences of metabolite profile between day shift work and night shift work in the case group
```{r}
require(nlme)
rst=NULL
for(i in valid_measures){
  metabo_aggre$m = metabo_aggre[,i]
  model = lme(m ~ shift + Alter + BMI + as.factor(AR_Rauch_zurzt) + as.factor(SD),
              metabo_aggre,
                subset = Kontrolle=="Night" & SW_Nr!="SW1041",
                random = ~ 1|SW_Nr,
                na.action=na.omit
  )
  rst = rbind(rst, summary(model)$tTable[2,])
}
rownames(rst) = valid_measures
rst = data.frame(rst)
rst$Value = -rst$Value
rst = data.frame(rst, fdr = p.adjust(rst$p.value, method = "BH"), bonf = p.adjust(rst$p.value, method = "bonf"))
write.csv(rst, file = "TEMP/categorized time_3h_Short term effect of night shift_mixed model_age_BMI_smoking_disease.csv")
```

#### 4.2.4 Plot of the metabolite concentrations in different sampling time intervals
```{r fig.height=5, fig.width=9, warning=FALSE}
require(ggplot2)
require(grid)

pdf("TEMP/metabolite concentration comparing day and night shift_both group_over 45.pdf", width = 30, height = 20)
k=1;nrow=1;ncol=1;newp=F
pushViewport(viewport(layout = grid.layout(3, 3)))
for(i in valid_measures){
  metabo_aggre$m = metabo_aggre[,i]
  
#   p = ggplot(metabo_aggre, aes(hour2, m))
#   p = p + geom_boxplot( aes(fill=interaction(factor(date2), factor(Kontrolle)) ) ) + ggtitle(i)
#   p = p + facet_grid(shift~.)
  
  tmp = metabo_aggre[which(metabo_aggre$Kontrolle=="Night"),]
  p = ggplot(tmp, aes(hour2, m))
  p = p + geom_boxplot( aes(fill=interaction(factor(date2), factor(shift)) ) ) + ggtitle(i)

  print(p, vp = viewport(layout.pos.row = nrow, layout.pos.col = ncol))
  if(ncol==3 & nrow==3) {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(3, 3)))
  }
  if(k<9) {
    if(k%%3==0) nrow=nrow+1
    k = k+1
    ncol = k-3*(nrow-1)
  }
  else {
    k=1
    nrow = 1
    ncol = 1
  }
    
}
dev.off()

p = ggplot(match_metabo_hormone, aes(x = Cortisol, y = C0, group = shift, col =factor(shift)))
p + geom_point() + geom_smooth()# + geom_line(aes(y = Cortisol), col="red")
```

#### 4.2.5 Plot correlation between metabolites and Cortisol
```{r warning=FALSE}
pdf("TEMP/metabolite correlation with Cortisol.pdf", width = 20, height = 20)
k=1;nrow=1;ncol=1;newp=F
pushViewport(viewport(layout = grid.layout(3, 3)))
for(i in valid_measures){
  match_metabo_hormone$m = match_metabo_hormone[,i]
  
  ##change the code here if plot some thing else
  p = ggplot(match_metabo_hormone, aes(x = Cortisol, y = m))
  p = p + geom_point() + geom_smooth(method = "lm")+ggtitle(i)
  ##
  
  print(p, vp = viewport(layout.pos.row = nrow, layout.pos.col = ncol))
  if(ncol==3 & nrow==3) {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(3, 3)))
  }
  if(k<9) {
    if(k%%3==0) nrow=nrow+1
    k = k+1
    ncol = k-3*(nrow-1)
  }
  else {
    k=1
    nrow = 1
    ncol = 1
  }
  
}
dev.off()
```
